---
title: "Recurrent Disease Outbreak in a Warm Temperate Marginal Coral Community"
author: "Joshua Heitzman"
date: "28 January 2021"
output:
  html_document:
    df_print: paged
  html_notebook: null
  pdf_document: default
  word_document: default
---
```{r setup, include=TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE,
	fig.path = "figures/",
	message = FALSE,
	warning = FALSE,
	dev = c("png", "pdf", "ps"))

library("scales")
library("ggplot2")
library("forcats")
library("tibble")
library("ggpubr")
library("patchwork")
library("dplyr")
library("tidyverse")
library("heatwaveR") # For detecting MHWs
library("phyloseq") #For ASV input
library("RColorBrewer")

Sys.setlocale("LC_ALL","English")
```

## Figure 1 - 2019 Disease Incidence (weekly surveys)
## Includes heatwave data with incidence from May 2019 to January 2020
```{R 2019 Disease Incidence, fig.width= 7, fig.height= 4}
#Setting up Heatwave data
OISST <- readRDS("shimoda_time_series_updated.rda")

# Detect the events in a time series
ts <- ts2clm(OISST, climatologyPeriod = c("1982-01-01", "2011-12-31"))
mhw <- detect_event(ts)

mhw2 <- mhw$climatology %>%
        filter(between(t, as.Date("2019-05-24"), as.Date("2020-02-10")))

mhw2.plot <- ggplot(data = mhw2, aes(x = t)) +

  geom_line(aes(y = temp, colour = "temp")) +
    #geom_line(aes(y = thresh, colour = "thresh"), size = 1, lty = "solid") +
    geom_line(aes(y = seas, colour = "seas"), size = 0.5, lty = "dotted") +
    #geom_flame(aes(y = temp, y2 = thresh, fill = "event"), show.legend = T) +
  scale_colour_manual(name = "Line Colour",
                      values = c("temp" = "black", 
                                 "thresh" =  "grey60", 
                                 "seas" = "black")) +
  scale_fill_manual(name = "Event Colour", 
                    values = c("event" = "red")) +
  scale_x_date(date_labels = "%b %Y") +
  guides(colour = guide_legend(override.aes = list(fill = NA))) +
  labs(y= expression(atop(paste(bold("Daily SST")), paste("(",degree ~C,")")))) +
  coord_cartesian(ylim = c(15, 30)) +
  theme_pubr() + theme(legend.position = "none") +
    theme(
      axis.title.x = element_blank(),
      axis.title.y = element_text (size = 9),
      axis.text.x = element_blank(),
      axis.text.y = element_text(size = 8),
      axis.ticks.x = element_blank()) +
    scale_x_date(limits = as.Date(c("2019-05-24", "2020-02-10")), expand = c(0,0), date_breaks = "1 month")

incidence.shidagaura.data <- read_csv("2019_Incidence_2.5.recheck.csv") %>% drop_na()

incidence.shidagaura.plot <- ggplot(data = incidence.shidagaura.data, aes(x = as.Date(Date), y = Incidence)) + 
  geom_point() +
  geom_segment(aes(xend = as.Date(Date), yend = 0 )) +
  labs(y = expression(atop(paste(bold("Disease Incidence")), paste("(New Occurrences)"))))+
  coord_cartesian(ylim = c(0, 30)) +
  geom_hline(yintercept = 0, color = "grey") +  
  scale_x_date(limits = as.Date(c("2019-05-24", "2020-02-10")), expand = c(0.004,0), date_breaks = "1 month", date_labels = "%b %Y") + 
  
theme_pubr() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_text (size = 10),
    axis.text.x = element_text (size  = 8),
    axis.text.y = element_text(size = 8))

mhw2.plot/incidence.shidagaura.plot + plot_annotation(tag_levels = 'A')
ggsave(filename = "Fig_1_Incidence.pdf", height = 3, width = 6)
```

##Figure 4 - ASV Table with bacteria taxon abundance from the WMS microbial mat compared with
##Healthy P. heronensis, sediment-trapping G. elegans and sediment
```{r ASV Input}

ASV_list <- read.csv("4_ASV_Percent.csv")
Taxa_list <- read.csv("Four_taxa.csv")

ASV <- otu_table(ASV_list, taxa_are_rows = TRUE)
TAX <- tax_table(as.matrix(Taxa_list))

physeq_all = phyloseq(ASV, TAX)
physeq_all_G = tax_glom(physeq_all, "Genus")
physeq_all_G_heatmap <- psmelt(physeq_all_G)
physeq_all_G_heatmap2 <- physeq_all_G_heatmap

physeq_all_G_heatmap2$Sample <- factor(physeq_all_G_heatmap2$Sample, 
  levels = c("ï..Disease", "Gelidium", "Health", "Sediment"), 
  labels = c("a","b", "c", "d"), ordered=TRUE)

physeq_names <- 
  physeq_all_G_heatmap2 %>% 
  mutate(Group = paste(as.character(Family), as.character(Genus), sep = " - ")) %>%
  pivot_wider(values_from = Abundance, names_from = Sample) %>%
  select(c("Group", "a","b", "c", "d")) %>%
  arrange(-a,-b,-c,-d)

physeq_all_G_heatmap$Sample <- factor(physeq_all_G_heatmap$Sample, 
  levels = c("ï..Disease", "Gelidium", "Health", "Sediment"), 
  labels = c("WMS","G. Elegans", "Healthy Coral", "Sediment"), ordered=TRUE)

physeq_all_G_heatmap_edit <- physeq_all_G_heatmap %>%
  mutate(Group=paste(Family, Genus, sep = " - ")) %>%
  mutate(Group=factor(Group,levels=rev(physeq_names$Group))) %>%
  mutate(Abundfactor=cut(Abundance,breaks=c(-1,0,0.001,0.01,0.05,0.1,0.25,0.5, 1),
    labels=c("0","0-0.1","0.1-1","1-5","5-10","10-25","25-50","50-100"))) %>%
  mutate(Abundfactor=factor(as.character(Abundfactor),levels=rev(levels(Abundfactor)))) %>%
  na.omit()
```

```{r heatmap, fig.width = 8, fig.height = 4}
textcol <- "grey40"

top_WMS <- physeq_all_G_heatmap_edit %>%
  filter(Sample == "WMS") %>%
  head(., n = 5) 


physeq_all_G_heatmap_edit%>%
    left_join(top_WMS,. , by = "Group") %>%

  ggplot( mapping = aes(x = Sample.y,y = Group,fill = Abundfactor.y)) +
  geom_tile() +
  xlab(label = "") + 
  theme_bw() +
  scale_fill_manual(values=rev(brewer.pal(8,"Blues")),na.value="grey90") +
  scale_y_discrete(expand=c(0,0)) +
  scale_x_discrete(expand=c(0,0)) +
  theme(legend.position="right",legend.direction="vertical",
        legend.title=element_blank(),
        legend.margin=margin(grid::unit(0,"cm")),
        legend.text=element_text(colour=textcol,size=12,face="bold"),
        legend.key.height=grid::unit(0.8,"cm"),
        legend.key.width=grid::unit(0.2,"cm"),
        axis.text.x=element_text(size=10,colour=textcol),
        axis.text.y=element_text(size=14, vjust=0.2,colour=textcol),
        axis.ticks=element_line(size=0.4),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        plot.background=element_blank(),
        panel.border=element_blank(),
        plot.margin=margin(0.7,0.4,0.1,0.2,"cm"),
        plot.title=element_text(colour=textcol,hjust=0,size=14,face="bold"))

heatmap.plot 
ggsave("heatmap.pdf",width = 8, height = 3)
```

## Supplementary Figure - Size Frequency of disease occurrences from 2019 disease incidence
```{r size frequency}
size.freq.data.2019 <- read.csv("2019_Histogram_2.5.recheck.csv") %>% drop_na() 
glimpse(size.freq.data.2019)
size.freq.data.2019$Size <- as.factor(size.freq.data.2019$Size)
size.freq.plot.2019 <- ggplot(data = size.freq.data.2019) +
  geom_col(aes(x = Size, y = Percent*100), position='dodge') +
  theme_pubr() +
    theme(
      axis.title.x = element_text (size = 10),
      axis.title.y = element_text (size = 10),
      axis.text.x = element_text(size = 8),
      axis.text.y = element_text(size = 8)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,100), labels = function(x) paste0(x, "%")) +
  coord_cartesian(ylim = c(0,50)) +
  labs(x = expression("Size (cm"^2*")"), y = "Size Frequency (%)")+
  theme(legend.position = "none")
  
size.freq.plot.2019
```

##Supplementary Figure
##2019 disease incidence of both sites (Shidagaura and Suzaki) paired with daily SST
```{r 2019 Disease Incidence (both sites), fig.height = 4, fig.width = 8}

# Detect the events in a time series
OISST <- readRDS("shimoda_time_series_updated.rda")
ts <- ts2clm(OISST, climatologyPeriod = c("1982-01-01", "2011-12-31"))
mhw <- detect_event(ts)

mhw2 <- mhw$climatology %>%
        filter(between(t, as.Date("2019-05-24"), as.Date("2020-02-15")))

mhw2.plot <- ggplot(data = mhw2, aes(x = t)) +

  geom_line(aes(y = temp, colour = "temp")) +
    #geom_line(aes(y = thresh, colour = "thresh"), size = 0.5, lty = "solid") +
    geom_line(aes(y = seas, colour = "seas"), size = 0.5, lty = "dotted") +
    #geom_flame(aes(y = temp, y2 = thresh, fill = "event"), show.legend = T) +
  scale_colour_manual(name = "Line Colour",
                      values = c("temp" = "black", 
                                 "thresh" =  "grey60", 
                                 "seas" = "black")) +
  scale_fill_manual(name = "Event Colour", 
                    values = c("event" = "red")) +
  scale_x_date(expand = c(0.004,0), date_breaks = "1 month", date_labels = "%b %Y") +
  guides(colour = guide_legend(override.aes = list(fill = NA))) +
  labs(y= expression(atop(paste(bold("Daily SST")), paste("( ",degree ~C," )"))))  +
  coord_cartesian(xlim = c(as.Date("2019-05-24"), as.Date("2020-02-15")), ylim = c(15, 30)) +
  theme_pubr() + theme(legend.position = "none") +
    theme(
      axis.title.x = element_blank(),
      axis.title.y = element_text (size = 9),
      axis.text.x = element_blank(),
      axis.text.y = element_text(size = 8),
      axis.ticks.x = element_blank(),
      plot.margin = margin(6,6,3,6, unit = "pt")) +
      scale_y_continuous(expand = c(0,0), limits = c(-1,30))

weekly.suzaki.data <- read_csv("2019_Incidence_2.5_Suzaki.csv") %>% 
  mutate(Date = as.Date(Date))  %>%
        filter(between(Date, as.Date("2019-05-24"), as.Date("2020-02-15")))

weekly.shidagaura.data <- read_csv("2019_Incidence_2.5.recheck.csv")%>% 
  mutate(Date = as.Date(Date))  %>%
        filter(between(Date, as.Date("2019-05-24"), as.Date("2020-02-15")))

weekly.suzaki.plot <- ggplot(data = weekly.suzaki.data, aes(x = Date, y = Incidence)) + 
  geom_point() +
  geom_segment(aes(xend = Date, yend = 0 )) +
  labs(x="",y = expression(atop(paste(bold("Site 2")), paste("Disease Incidence")))) +
  geom_hline(yintercept = 0, color = "grey") +
  theme_pubr() +
  theme(
      axis.title.x = element_text (size = 10),
      axis.title.y = element_text (size = 10),
      axis.text.x = element_text (size = 8),
      axis.text.y = element_text(size = 8)) +
  scale_x_date(expand = c(0.004,0),date_breaks = "1 month", date_labels = "%b %Y") +
  scale_y_continuous(expand = c(0,0), limits = c(-1,30)) +
    coord_cartesian(xlim = c(as.Date("2019-05-24"), as.Date("2020-02-15")))

weekly.shidagaura.plot <- ggplot(data = weekly.shidagaura.data, aes(x = Date, y = Incidence)) + 
  geom_point() +
  geom_segment(aes(xend = Date, yend = 0 )) +
  labs(x="",y = expression(atop(paste(bold("Site 1")), paste("Disease Incidence")))) +
  geom_hline(yintercept = 0, color = "grey") +
  theme_pubr() +
  theme(
axis.title.x = element_blank(),
      axis.title.y = element_text (size = 10),
      axis.text.x = element_blank(),
      axis.text.y = element_text(size = 8),
      axis.ticks.x = element_blank())+
      scale_x_date(expand = c(0.004,0)) +
      scale_y_continuous(expand = c(0,0), limits = c(-1,30))+
    coord_cartesian(xlim = c(as.Date("2019-05-24"), as.Date("2020-02-15")))

mhw2.plot/weekly.shidagaura.plot/weekly.suzaki.plot + plot_annotation(tag_levels = "A")
ggsave(filename = "Fig_Supp_Incidence.pdf", height = 5, width = 6)
```

## Disease Prevalence from September 2013 to December 2020
## Paired with daily SST of the area and heatwave occurrence
```{r fig 2 Prevalence}
OISST <- readRDS("shimoda_time_series_updated.rda")

ts <- ts2clm(OISST, climatologyPeriod = c("1982-01-01", "2011-12-31"))
mhw <- detect_event(ts)

mhw2 <- mhw$climatology %>%
        filter(between(t, as.Date("2013-09-13"), as.Date("2020-12-08")))

mhw2.plot <- ggplot(data = mhw2, aes(x = t)) +
  geom_line(aes(y = temp, colour = "temp"), size = 0.5) +
#  geom_flame(aes(y = temp, y2 = thresh, fill = "event"), show.legend = T) +
  scale_colour_manual(name = "Line Colour",
                      values = c("temp" = "black", 
                                 "thresh" =  "grey60", 
                                 "seas" = "black")) +
  scale_fill_manual(name = "Event Colour", 
                    values = c("event" = "red")) +
  scale_x_date(expand = c(0.004,0.004), date_breaks = "1 year", date_labels = "%b %Y") +
  guides(colour = guide_legend(override.aes = list(fill = NA))) +
  labs(y= expression(atop(paste(bold("Daily SST")), paste("(",degree ~C," )")))) +
  coord_cartesian(ylim = c(10, 30)) +
  theme_pubr() + theme(legend.position = "none") +
    theme(
      axis.title.x = element_blank(),
      axis.title.y = element_text (size = 8),
      axis.text.x = element_blank(),
      axis.text.y = element_text(size = 7))

#Setting up prevalence data
#Prevalence of sediment-trapping G. elegans, prevalence of disease and number of disease occurrences per survey
SC.G.elegans.data <- 
  read_csv("Dis_Monthly_Monitoring.csv") %>%
  mutate(Date = as.Date(Date)) %>%
  mutate(Coverage_Deg_G_elegans = (Deg_G_Elegans/64)*100) %>% 
  group_by(Date) %>%
  get_summary_stats(type = "mean_sd",WMS_Prevalence, Coverage_Deg_G_elegans) %>%
  filter(variable %in% "Coverage_Deg_G_elegans")

Prevalence.data <- 
  read_csv("Dis_Monthly_Monitoring.csv") %>%
  mutate(Date = as.Date(Date)) %>%
  filter(between(Date, as.Date("2013-09-13"), as.Date("2020-12-08"))) %>%
  mutate(Coverage_Deg_G_elegans = (Deg_G_Elegans/64)*100) %>%
  mutate(Percent_Prev = (WMS_Prevalence*100)) %>%
  group_by(Date) %>%  
  get_summary_stats(type = "mean_sd",Percent_Prev, Coverage_Deg_G_elegans) %>%
  filter(variable %in% "Percent_Prev")

SC.GE.plot <-
  ggplot(data = SC.G.elegans.data, aes(x = Date)) +
  geom_line(aes(y = mean)) +
  geom_point(aes(y = mean),size = 1) +
  geom_errorbar(aes(ymax = mean + sd, ymin = mean - sd)) +
  labs(y= expression(atop(paste(bold("Sediment Trapping"~bolditalic("G. elegans"))), paste("(% of Total Coverage)")))) +
  geom_hline(yintercept = 0, color = "grey") +
  theme_pubr() +
    theme(
      axis.title.x = element_blank(),
      axis.title.y = element_text(size = 8),
      axis.text.x = element_text(size = 8),  
      axis.text.y = element_text(size = 7)) +  
  coord_cartesian(ylim = c(0, 20)) +
  scale_x_date(expand = c(0.004,0.004), date_breaks = "1 year", date_labels = "%Y")

Dis.Prev.plot <-
  ggplot(data = Prevalence.data, aes(x = Date, y = mean)) +
  geom_line() +
  geom_point(size = 1) +
  geom_errorbar(aes(ymax = mean + sd, ymin = mean - sd)) +
  labs(y= expression(atop(paste(bold("Disease")), paste("(% of Coral Coverage)")))) +
  geom_hline(yintercept = 0, color = "grey") +
  theme_pubr() +
    theme(
      axis.title.x = element_blank(),
      axis.title.y = element_text(size = 8),
      axis.text.x = element_blank(),
      axis.text.y = element_text(size = 7) )+
  coord_cartesian(ylim = c(0, 40)) +
  scale_x_date(expand = c(0.004,0.004), date_breaks = "1 year", date_labels = "%Y")  

Dis_Count.data <- read_csv("Dis_Count.csv")
Dis_Count.plot <- ggplot(data = Dis_Count.data, aes(x = Date, y = Dis_Count)) + 
  geom_point(size = 1) +
  geom_line() +
  labs(x = "", y= expression(atop(paste(bold("Disease")), paste("(# of Occurrences)")))) +
  geom_hline(yintercept = 0, color = "grey") +
  theme_pubr() +
    theme(
      axis.title.x = element_blank(),
      axis.title.y = element_text(size = 8),
      axis.text.x = element_blank(),
      axis.text.y = element_text(size = 7) )+
  scale_x_date(expand = c(0.004,0.004), date_breaks = "1 year", date_labels = "%Y")

Dis.Prev.plot/Dis_Count.plot/mhw2.plot/SC.GE.plot + plot_annotation(tag_levels = 'A')
```

##Calculation of average temperature during disease outbreaks
##2013 omitted due to surveys beginning at the end of the year (September 2013)
```{r Temp Calc}
OISST <- readRDS("shimoda_time_series_updated.rda")

ts <- ts2clm(OISST, climatologyPeriod = c("1982-01-01", "2011-12-31"))
mhw <- detect_event(ts)

mhw2014 <- mhw$climatology %>%
        filter(between(t, as.Date("2014-06-16"), as.Date("2014-11-10")))
mhw2015 <- mhw$climatology %>%
        filter(between(t, as.Date("2015-09-23"), as.Date("2015-11-24")))
mhw2016 <- mhw$climatology %>%
        filter(between(t, as.Date("2016-09-05"), as.Date("2016-12-20")))
mhw2017 <- mhw$climatology %>%
        filter(between(t, as.Date("2017-08-25"), as.Date("2017-12-11")))
mhw2018 <- mhw$climatology %>%
        filter(between(t, as.Date("2018-08-18"), as.Date("2018-11-02")))
mhw2019 <- mhw$climatology %>%
        filter(between(t, as.Date("2019-09-02"), as.Date("2019-11-15")))
mhw2020 <- mhw$climatology %>%
        filter(between(t, as.Date("2020-09-15"), as.Date("2020-10-14")))

mhw_outbreak <- bind_rows(mhw2014,mhw2015,mhw2016,mhw2017,mhw2018,mhw2019,mhw2020) %>% 
  get_summary_stats(temp,type = "mean_sd")
```

